Please set the folder "Tingting Sun" as the working directory before run this file
Make sure the alldata.xlsx is not changed
```{r}
library(openxlsx)
wd=getwd()
fullPath=paste0(wd,"/PA/alldata.xlsx")
alldata=read.xlsx(fullPath)
x = alldata
x[,3] = as.Date(x[,3],origin="1899-12-30")
x[,4] = as.Date(x[,4],origin="1899-12-30")
x[,15] = as.Date(x[,15],origin = "1899-12-30")
gapdays = c()
for(i in 2:length(x$`PATIENT/ENCOUNTER.NUMBER`)){
  if(x$PATIENT.HISTORY.NUMBER[i] == x$PATIENT.HISTORY.NUMBER[i-1]){
  gapdays[i-1] = abs(x$PATIENT.ADMIT.DATE[i]-x$PATIENT.DISCHARGE.DATE[i-1])
  }else{
  gapdays[i-1] = NA
  }
}
gapdays[length(x$`PATIENT/ENCOUNTER.NUMBER`)] = -1
x=cbind(x,gapdays)
index = which(gapdays >-1)
x = x[index,]
index = which(x$Patient.Days>0)
x = x[index,]

#age Transform age to age group, 0-10, 10-20, 20-30, …. Then delete the date of birth.
age = c()
for(i in 1:length(x$`PATIENT/ENCOUNTER.NUMBER`)){
  age[i] = floor((x$PATIENT.ADMIT.DATE[i]-x$PATIENT.DATE.OF.BIRTH[i])/3650)
}
x$Age = age
x = x[,-15]

#Nur.Stat
x$Nur.Stat[which(is.na(x$Nur.Stat))] = 'isNA'

#Transform zipcode to low-income, middle-income, high-income.
library(openxlsx)
fullPath=paste0(wd,"/data/income_zip.xlsx")
zip = read.xlsx(fullPath)
colnames(x)[18] = 'ZipCode'
library("data.table")
dat = x
dat = data.table(dat)
dat$ZipCode = as.character(dat$ZipCode)
zip = data.table(zip)
zip$ZipCode = as.character(zip$ZipCode)
zip = zip[,1:2]
library("dplyr")
x = left_join(dat, zip,by = "ZipCode")
label = 1:10
x[,22]=cut(as.numeric(x[,22]),breaks=c(10000,20000,30000,40000,50000,60000,70000,80000,90000,100000,110000),label,ordered_result = T)
colnames(x)[22] = 'income_level' 
x <- x[c(1:(ncol(x)-2),22,21)]
x = na.omit(x)
#write.xlsx(x,file="C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/data.xlsx")
```

#Task1–Explore the relationship of each variable to gapdays
```{r}
library(openxlsx)
#data = read.xlsx("C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/data.xlsx")
data = x
for( i in 5:22){
  if(i %in% c(5, 6, 22)){
    data[,i]=as.numeric(data[,i])
  }else{
    data[,i]=as.factor(data[,i])
  }
}
library(ggplot2)
p <- ggplot(data, aes(x = log(gapdays + 1))) + geom_histogram()
print(p)

vars <- colnames(data)[7:ncol(data)-1]
for (i in vars) {
  plot <- ggplot(data, aes(x=as.factor(data[,i]),y=log(gapdays))) + geom_boxplot() + labs(x=i)
  print(plot)
}
```

#Task2–Reduce the number of factor levels where appropriate
```{r}
library(plyr)
library(dplyr)
#dat = read.xlsx("C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/data.xlsx")
dat = x

dat = plyr::rename(dat, c("PATIENT/ENCOUNTER.NUMBER"="PatientEncounterNumber","PATIENT.HISTORY.NUMBER"="PatientHistoryNumber","PATIENT.HISTORY.NUMBER"="PatientHistoryNumber","PATIENT.DISCHARGE.DATE"="PatientDischargeDate","I/O.CODE"="IO_CODE","ICD-9.DIAGNOST.CODE"="ICD9_DIAGNOST_CODE"))
vars <- colnames(dat)[6:length(dat)-1]
for (i in vars) {
  plot <- ggplot(dat, aes(x=dat[,i])) + geom_bar() + labs(x=i) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
  print(plot)
}
for( i in 5:22){
  if(i %in% c(5, 6, 22)){
    dat[,i]=as.numeric(dat[,i])
  }else{
    dat[,i]=as.factor(dat[,i])
  }
}

k = kmeans(dat$DOCTOR.NUMBER,10)
dat$DOCTOR.NUMBER = k$cluster
k = kmeans(dat$PATIENT.DRG,10)
dat$PATIENT.DRG = k$cluster
k = kmeans(dat$Surgeon,10)
dat$Surgeon = k$cluster
k = kmeans(dat$DISCHARGE.STATUS,10)
dat$DISCHARGE.STATUS = k$cluster
k = kmeans(dat$ICD.PROCEDURE.CODE,10)
dat$ICD.PROCEDURE.CODE = k$cluster
k = kmeans(dat$ICD9_DIAGNOST_CODE,10)
dat$ICD9_DIAGNOST_CODE = k$cluster
#write.xlsx(x,file="C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/dat.xlsx")
```

```{r}
vars<-colnames(dat)[7:ncol(dat)-1]
for (i in vars) {
  print(i)
  x <- dat %>% group_by_(i)%>%dplyr::summarise(mean = mean(log(gapdays+1)),median = median(log(gapdays+1))              ,n = n())#dplyr中的summarise跟plyr中的不一样 
  print(x)
}
```
# PATIENT.DRG
#change1
```{r}
#PATIENT.DRG
i="PATIENT.DRG"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$PATIENT.DRG <- as.factor(dat$PATIENT.DRG)
levels(dat[,"PATIENT.DRG"])
dat2 <- dat
var <- "PATIENT.DRG"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("1528","1528","310","49","1528","6","7","1528","49","310"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])


```

#Nur.Stat
# Change 2: 
```{r}
#Nur.Stat
i="Nur.Stat"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$Nur.Stat <- as.factor(dat$Nur.Stat)
levels(dat[,"Nur.Stat"])
dat2 <- dat
var <- "Nur.Stat"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("MS_CCU_ER_PC", "MS_CCU_ER_PC","isNA", "MS_CCU_ER_PC",  "MS_CCU_ER_PC",  "MS_CCU_ER_PC",  "WS"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])

```

#DOCTOR.NUMBER
#change3 
```{r}
#DOCTOR.NUMBER
i="DOCTOR.NUMBER"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$DOCTOR.NUMBER <- as.factor(dat$DOCTOR.NUMBER)
levels(dat[,"DOCTOR.NUMBER"])
dat2 <- dat
var <- "DOCTOR.NUMBER"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("r","r","r","r","r","610","r","r","r","r"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])

```

#Surgeon
#change4
```{r}
#Surgeon
i="Surgeon"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$Surgeon <- as.factor(dat$Surgeon)
levels(dat[,"Surgeon"])
dat2 <- dat
var <- "Surgeon"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("1257","1257","346","346","1257","346","1257","8","9","10"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])

```
#HOSPITAL.SERVICE.CODE
# Change 5
```{r}
#HOSPITAL.SERVICE.CODE
i="HOSPITAL.SERVICE.CODE"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$HOSPITAL.SERVICE.CODE <- as.factor(dat$HOSPITAL.SERVICE.CODE)
levels(dat[,"HOSPITAL.SERVICE.CODE"])
dat2 <- dat
var <- "HOSPITAL.SERVICE.CODE"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("R","R","R","R","581415","R","R","581415","R","R","R","R","R","581415","581415"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
max <- which.max(table[,2])
level.name <- as.character(table[max,1])
dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
```

#ADMIT.SOURCE
#change6
```{r}
#ADMIT.SOURCE
i="ADMIT.SOURCE"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$ADMIT.SOURCE <- as.factor(dat$ADMIT.SOURCE)
levels(dat[,"ADMIT.SOURCE"])
dat2 <- dat
var <- "ADMIT.SOURCE"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("156","234","234","234","156","156","7"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
```

#DISCHARGE.STATUS
#change7 ??
```{r}
#DISCHARGE.STATUS
i="DISCHARGE.STATUS"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$DISCHARGE.STATUS <- as.factor(dat$DISCHARGE.STATUS)
levels(dat[,"DISCHARGE.STATUS"])
dat2 <- dat
var <- "DISCHARGE.STATUS"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("12359","12359","12359","467810","12359","467810","467810","467810","12359","467810"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
```
#PATIENT.SEX.CODE
#change8
```{r}
#PATIENT.SEX.CODE
dat$PATIENT.SEX.CODE <- as.factor(dat$PATIENT.SEX.CODE)
levels(dat[,"PATIENT.SEX.CODE"])
dat2 <- dat
var <- "PATIENT.SEX.CODE"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("F","M","F"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
```
#PATIENT.MARITAL.STATUS
#change9
```{r}
#PATIENT.MARITAL.STATUS
i="PATIENT.MARITAL.STATUS"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$PATIENT.MARITAL.STATUS <- as.factor(dat$PATIENT.MARITAL.STATUS)
levels(dat[,"PATIENT.MARITAL.STATUS"])
dat2 <- dat
var <- "PATIENT.MARITAL.STATUS"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
var.levels
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("DWX","MPS","MPS","MPS","U","DWX","DWX"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
```
#ZipCode
#change10
#ICD.PROCEDURE.CODE
#change11
```{r}
#ICD.PROCEDURE.CODE
i="ICD.PROCEDURE.CODE"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$ICD.PROCEDURE.CODE <- as.factor(dat$ICD.PROCEDURE.CODE)
levels(dat[,"ICD.PROCEDURE.CODE"])
dat2 <- dat
var <- "ICD.PROCEDURE.CODE"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("138","2","138","45710","138","6","45710","138","9","45710"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
```
#ICD9_DIAGNOST_CODE
#change12
```{r}
#ICD9_DIAGNOST_CODE
i="ICD9_DIAGNOST_CODE"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$ICD9_DIAGNOST_CODE <- as.factor(dat$ICD9_DIAGNOST_CODE)
levels(dat[,"ICD9_DIAGNOST_CODE"])
dat2 <- dat
var <- "ICD9_DIAGNOST_CODE"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("1","27","3469","3469","510","3469","27","8","3469","510"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
```

#income_level
#change13
```{r}
#income_level
i="income_level"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$income_level <- as.factor(dat$income_level)
levels(dat[,"income_level"])
dat2 <- dat
var <- "income_level"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("12","r","12","r","r","r","r","r","r","r"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
```
```{r}
dat = dat2
#write.xlsx(dat,file="C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/dat.xlsx")
```
#Task3–Use observations from principal components analysis (PCA) to generate a new feature
```{r}
library(openxlsx)
#dat = read.xlsx("C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/dat.xlsx")
vars <- colnames(dat)[6:length(dat)-1]

for( i in 5:22){
  if(i %in% c( 6, 22)){
    dat[,i]=as.numeric(dat[,i])
  }else{
    dat[,i]=as.factor(dat[,i])
  }
}
#These predictors could be correlated: PATIENT.DRG and HOSPITAL.SERVICE.CODE and ICD9.PROCEDURE.CODE and ICD9_DIAGNOST_CODE could be related.
#So I explore a PCA using these predictors.
x_list=c("PATIENT.DRG","HOSPITAL.SERVICE.CODE","ICD.PROCEDURE.CODE","ICD9_DIAGNOST_CODE")
datPCA = dat[,x_list]
library(caret)
varsPCA = colnames(datPCA)
for (var in varsPCA) {
  datPCA[, var] <- as.character(datPCA[, var])
}
binarizer <- caret::dummyVars(paste("~", paste(varsPCA, collapse = "+")) , data = datPCA, fullRank = FALSE)
datPCAbin <- data.frame(predict(binarizer, newdata = datPCA))
head(datPCAbin)
PCAweather <- prcomp(datPCAbin, center = TRUE, scale. = TRUE)
summary(PCAweather)
resultPCA <- PCAweather$rotation
```



```{r}

```
```{r}

```
```{r}

```
```{r}

```
```{r}

```
