Please set the folder "Tingting Sun" as the working directory before run this file
Make sure the alldata.xlsx is not changed

1. relevel PATIENT.MARITAL.STATUS;HOSPITAL.SERVICE.CODE;DISCHARGE.STATUS
2. rename age
3. change the income level from 10levels to 3levels (low medium high)
4.Patient Days less than or equal to 19 are combined as one level called 0-19. Patient Days from 20 to 30 are grouped to the level called 20-30. The others are grouped to 30+.
5.delete first 4 variables
6. I have different PCAresults 
7. 
```{r}
library(openxlsx)
#wd=getwd()
#fullPath=paste0(wd,"/PA/alldata.xlsx")
#alldata = read.xlsx("C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/alldata.xlsx")
if(Sys.info()["user"]=="luxiong"){
  alldata <- read.xlsx("/Users/luxiong/Dropbox/Actuarial research seminar/Humana case competition/Team folder/Tingting Sun/PA/alldata.xlsx")
}else{
  alldata <- read.xlsx("C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/alldata.xlsx")
}
#alldata=read.xlsx(fullPath)
x = alldata
x[,3] = as.Date(x[,3],origin="1899-12-30")
x[,4] = as.Date(x[,4],origin="1899-12-30")
x[,15] = as.Date(x[,15],origin = "1899-12-30")
gapdays = c()
for(i in 2:length(x$`PATIENT/ENCOUNTER.NUMBER`)){
  if(x$PATIENT.HISTORY.NUMBER[i] == x$PATIENT.HISTORY.NUMBER[i-1]){
  gapdays[i-1] = abs(x$PATIENT.ADMIT.DATE[i]-x$PATIENT.DISCHARGE.DATE[i-1])
  }else{
  gapdays[i-1] = NA
  }
}
gapdays[length(x$`PATIENT/ENCOUNTER.NUMBER`)] = -1
x=cbind(x,gapdays)
index = which(gapdays >-1)
x = x[index,]
index = which(x$Patient.Days>0)
x = x[index,]
#age Transform age to age group, 0-10, 10-20, 20-30, …. Then delete the date of birth.
age = c()
for(i in 1:length(x$`PATIENT/ENCOUNTER.NUMBER`)){
  age[i] = floor((x$PATIENT.ADMIT.DATE[i]-x$PATIENT.DATE.OF.BIRTH[i])/3650)
}
x$Age = age
x = x[,-15]
x$Age = factor(as.factor(x$Age),levels=c("0","1","2","3","4","5","6","7","8","9","10"),labels=c('0-10','10-20','20-30','30-40','40-50','50-60','60-70','70-80','80-90','90-100','100+'),ordered = TRUE)
#Nur.Stat
x$Nur.Stat[which(is.na(x$Nur.Stat))] = 'isNA'
#Transform zipcode to low-income, middle-income, high-income.
library(openxlsx)
if(Sys.info()["user"]=="luxiong"){
  zip = read.xlsx("/Users/luxiong/Dropbox/Actuarial research seminar/Humana case competition/Team folder/Tingting Sun/H2O case/data/income_zip.xlsx")
}else{
  zip = read.xlsx("C:/Users/HP/Dropbox/Team folder/Tingting Sun/H2O case/data/income_zip.xlsx")  
  }
colnames(x)[18] = 'ZipCode'
library("data.table")
dat = x
dat = data.table(dat)
dat$ZipCode = as.character(dat$ZipCode)
zip = data.table(zip)
zip$ZipCode = as.character(zip$ZipCode)
zip = zip[,1:2]
library("dplyr")
x = left_join(dat, zip,by = "ZipCode")
label = c("low","meduim","high")
x[,22]=cut(as.numeric(x[,22]),breaks=c(0,40000,50000,1100000),label,ordered_result = T)
colnames(x)[22] = 'income_level' 
x <- x[c(1:(ncol(x)-2),22,21)]
x = na.omit(x)
x = x[,-18] #delete ZipCode

#Patient Days less than or equal to 19 are combined as one level called 0-19. Patient Days from 20 to 30 are grouped to the level called 20-30. The others are grouped to 30+.
label = c("0-19","20-30","30+")
x[,6]=cut(as.numeric(x[,6]),breaks=c(-1,20,31,10000000),label,ordered_result = T)
index = 1:4
x = x[,-index]
#write.xlsx(x,file="C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/data.xlsx")
```

#Task1–Explore the relationship of each variable to gapdays
```{r}
library(openxlsx)
data=x
#if(Sys.info()["user"]=="luxiong")
#data = read.xlsx("C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/data.xlsx")
for( i in 1:17){
  if(i %in% c(17)){
    data[,i]=as.numeric(data[,i])
  }else{
    data[,i]=as.factor(data[,i])
  }
}
library(ggplot2)
p <- ggplot(data, aes(x = log(gapdays + 1))) + geom_histogram()
print(p)

vars <- colnames(data)[1:16]
for (i in vars) {
  data$Age = factor(as.factor(data$Age),levels=c('0-10','10-20','20-30','30-40','40-50','50-60','60-70','70-80','80-90','90-100','100+'),labels=c('0-10','10-20','20-30','30-40','40-50','50-60','60-70','70-80','80-90','90-100','100+'),ordered = TRUE)
  plot <- ggplot(data, aes(x=data[,i],y=log(gapdays))) + geom_boxplot() + labs(x=i)
  print(plot)
}
```
#Task2–Reduce the number of factor levels where appropriate
```{r}
library(plyr)
library(dplyr)
dat=x
#dat = read.xlsx("C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/data.xlsx")
dat = plyr::rename(dat, c("PATIENT/ENCOUNTER.NUMBER"="PatientEncounterNumber","PATIENT.HISTORY.NUMBER"="PatientHistoryNumber","PATIENT.HISTORY.NUMBER"="PatientHistoryNumber","PATIENT.DISCHARGE.DATE"="PatientDischargeDate","I/O.CODE"="IO_CODE","ICD-9.DIAGNOST.CODE"="ICD9_DIAGNOST_CODE"))
vars <- colnames(dat)[1:length(dat)-1]
for (i in vars) {
  plot <- ggplot(dat, aes(x=dat[,i])) + geom_bar() + labs(x=i) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
  print(plot)
}
for( i in 1:17){
  if(i %in% c(17)){
    dat[,i]=as.numeric(dat[,i])
  }else{
    dat[,i]=as.factor(dat[,i])
  }
}
 
#write.xlsx(x,file="C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/dat.xlsx")
```

```{r}
vars<-colnames(dat)[1:ncol(dat)-1]
for( i in 1:17){
  if(i %in% c(17)){
    dat[,i]=as.numeric(dat[,i])
  }else{
    dat[,i]=as.factor(dat[,i])
  }
}
for (i in vars) {
  options(digits = 6)
  print(i)
  data$Age = factor(as.factor(data$Age),levels=c('0-10','10-20','20-30','30-40','40-50','50-60','60-70','70-80','80-90','90-100','100+'),labels=c('0-10','10-20','20-30','30-40','40-50','50-60','60-70','70-80','80-90','90-100','100+'),ordered = TRUE)
  x <- dat %>% group_by_(i)%>%dplyr::summarise(mean = mean(log(gapdays+1)),median = median(log(gapdays+1))              ,n = n())#dplyr中的summarise跟plyr中的不一样 
  #print(x)
  x = as.data.frame(x)
  print(x[order(as.numeric(x[,2]),decreasing= T),])
}
```

# PATIENT.DRG
#change1
```{r}
dat3 = dat
dat2 = dat
#PATIENT.DRG
i="PATIENT.DRG"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$PATIENT.DRG <- as.factor(dat$PATIENT.DRG)
levels(dat[,"PATIENT.DRG"])
var <- "PATIENT.DRG"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c(as.factor(groups)))
#Relevel
table <- as.data.frame(table(dat2[,var]))
max <- which.max(table[,2])
level.name <- as.character(table[max,1])
dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
dat3$PATIENT.DRG = dat2$PATIENT.DRG
```

#Nur.Stat
# Change 2: 
```{r}
#Nur.Stat

dat$Nur.Stat <- as.factor(dat$Nur.Stat)
levels(dat[,"Nur.Stat"])
var <- "Nur.Stat"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("MS_CCU_ER_PC", "MS_CCU_ER_PC","isNA", "MS_CCU_ER_PC",  "MS_CCU_ER_PC",  "MS_CCU_ER_PC",  "WS"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
dat3$Nur.Stat = dat2$Nur.Stat
```

#DOCTOR.NUMBER
#change3  ???
```{r}
#DOCTOR.NUMBER
i="DOCTOR.NUMBER"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$DOCTOR.NUMBER <- as.factor(dat$DOCTOR.NUMBER)
levels(dat[,"DOCTOR.NUMBER"])
var <- "DOCTOR.NUMBER"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c(as.factor(groups)))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
dat3$DOCTOR.NUMBER = dat2$DOCTOR.NUMBER
```

#Surgeon
#change4 ????
```{r}
#Surgeon
i="Surgeon"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$Surgeon <- as.factor(dat$Surgeon)
levels(dat[,"Surgeon"])
var <- "Surgeon"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c(as.factor(groups)))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
dat3$Surgeon = dat2$Surgeon
```

#HOSPITAL.SERVICE.CODE
# Change 5
```{r}
#HOSPITAL.SERVICE.CODE
dat$HOSPITAL.SERVICE.CODE <- as.factor(dat$HOSPITAL.SERVICE.CODE)
levels(dat[,"HOSPITAL.SERVICE.CODE"])
var <- "HOSPITAL.SERVICE.CODE"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("Y","R","R","G","B","R","Y","B","R","R","Y","R","B","B","B"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
max <- which.max(table[,2])
level.name <- as.character(table[max,1])
dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
dat3$HOSPITAL.SERVICE.CODE = dat2$HOSPITAL.SERVICE.CODE
```

#ADMIT.SOURCE
#change6----delete
```{r}
#ADMIT.SOURCE
i="ADMIT.SOURCE"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$ADMIT.SOURCE <- as.factor(dat$ADMIT.SOURCE)
levels(dat[,"ADMIT.SOURCE"])
var <- "ADMIT.SOURCE"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c(as.factor(groups)))
#Relevel
table <- as.data.frame(table(dat2[,var]))
max <- which.max(table[,2])
level.name <- as.character(table[max,1])
dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
dat3$ADMIT.SOURCE = dat2$ADMIT.SOURCE
```

#DISCHARGE.STATUS
#change7 
```{r}
#DISCHARGE.STATUS
dat$DISCHARGE.STATUS <- as.factor(dat$DISCHARGE.STATUS)
levels(dat[,"DISCHARGE.STATUS"])
var <- "DISCHARGE.STATUS"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("R","Y","G","G","Y","G","R","G","G","R","Y","G","G","Y","Y","G","Y","Y","Y","G","G","G"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
dat3$DISCHARGE.STATUS = dat2$DISCHARGE.STATUS
```
#PATIENT.SEX.CODE
#change8
```{r}
#PATIENT.SEX.CODE
dat$PATIENT.SEX.CODE <- as.factor(dat$PATIENT.SEX.CODE)
levels(dat[,"PATIENT.SEX.CODE"])

var <- "PATIENT.SEX.CODE"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("F","M","F"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
  max <- which.max(table[,2])
  level.name <- as.character(table[max,1])
  dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
dat3$PATIENT.SEX.CODE = dat2$PATIENT.SEX.CODE
```
#PATIENT.MARITAL.STATUS
#change9
```{r}
#PATIENT.MARITAL.STATUS
dat$PATIENT.MARITAL.STATUS <- as.factor(dat$PATIENT.MARITAL.STATUS)
levels(dat[,"PATIENT.MARITAL.STATUS"])
var <- "PATIENT.MARITAL.STATUS"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("DUWXNA","M","PS","PS","DUWXNA","DUWXNA","DUWXNA"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
max <- which.max(table[,2])
level.name <- as.character(table[max,1])
dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
dat3$PATIENT.MARITAL.STATUS = dat2$PATIENT.MARITAL.STATUS
```
#change10
#PATIENT.RACE.CODE
```{r}
#PATIENT.RACE.CODE
dat$PATIENT.RACE.CODE
dat$PATIENT.RACE.CODE <- as.factor(dat$PATIENT.RACE.CODE)
levels(dat[,"PATIENT.RACE.CODE"])
var <- "PATIENT.RACE.CODE"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c("AIMNT","BO","DH","DH","AIMNT","AIMNT","AIMNT","BO","AIMNT","WX","WX"))
#Relevel
table <- as.data.frame(table(dat2[,var]))
max <- which.max(table[,2])
level.name <- as.character(table[max,1])
dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
dat3$PATIENT.RACE.CODE = dat2$PATIENT.RACE.CODE
```

#ICD.PROCEDURE.CODE
#change11
```{r}
#ICD.PROCEDURE.COD

i="ICD.PROCEDURE.CODE"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$ICD.PROCEDURE.CODE <- as.factor(dat$ICD.PROCEDURE.CODE)
levels(dat[,"ICD.PROCEDURE.CODE"])
var <- "ICD.PROCEDURE.CODE"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c(as.factor(groups)))
#Relevel
table <- as.data.frame(table(dat2[,var]))
max <- which.max(table[,2])
level.name <- as.character(table[max,1])
dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
dat3$ICD.PROCEDURE.CODE = dat2$ICD.PROCEDURE.CODE
```
#ICD9_DIAGNOST_CODE
#change12
```{r}
#ICD9_DIAGNOST_CODE
i="ICD9_DIAGNOST_CODE"
x=dat %>% group_by_(i)%>%dplyr::summarise(mean=mean(log(gapdays+1)),
                                   median=median(log(gapdays+1)),
                                   n = n())
#Reference: https://www.statmethods.net/advstats/cluster.html
mydata=x$mean
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

dat$ICD9_DIAGNOST_CODE <- as.factor(dat$ICD9_DIAGNOST_CODE)
levels(dat[,"ICD9_DIAGNOST_CODE"])
var <- "ICD9_DIAGNOST_CODE"
dat2[,var] <- as.factor(dat2[,var])
var.levels <- levels(dat2[,var])
dat2[,var] <- mapvalues(dat2[,var],var.levels,c(as.factor(groups)))
#Relevel
table <- as.data.frame(table(dat2[,var]))
max <- which.max(table[,2])
level.name <- as.character(table[max,1])
dat2[,var] <- relevel(dat2[,var], ref = level.name)
table(dat2[,var])
dat3$ICD9_DIAGNOST_CODE = dat2$ICD9_DIAGNOST_CODE
```

```{r}
dat = dat3
#write.xlsx(dat3,file="C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/dat3.xlsx")
```

#Task3–Use observations from principal components analysis (PCA) to generate a new feature
```{r}
library(openxlsx)
#dat = read.xlsx("C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/dat3.xlsx")
vars <- colnames(dat)[1:length(dat)-1]

for( i in 1:17){
  if(i %in% c(17)){
    dat[,i]=as.numeric(dat[,i])
  }else{
    dat[,i]=as.factor(dat[,i])
  }
}
#These predictors could be correlated: PATIENT.DRG and HOSPITAL.SERVICE.CODE and ICD9.PROCEDURE.CODE and ICD9_DIAGNOST_CODE could be related.
#So I explore a PCA using these predictors.
x_list=c("PATIENT.DRG","HOSPITAL.SERVICE.CODE","ICD.PROCEDURE.CODE","ICD9_DIAGNOST_CODE")
datPCA = dat[,x_list]
library(caret)
varsPCA = colnames(datPCA)
for (var in varsPCA) {
  datPCA[, var] <- as.character(datPCA[, var])
}
binarizer <- caret::dummyVars(paste("~", paste(varsPCA, collapse = "+")) , data = datPCA, fullRank = FALSE)
datPCAbin <- data.frame(predict(binarizer, newdata = datPCA))
head(datPCAbin)
PCAweather <- prcomp(datPCAbin, center = TRUE, scale. = TRUE)
summary(PCAweather)
phi <- PCAweather$rotation
phi <- as.data.frame(phi)
```

```{r}
c = cbind(row.names(phi),abs(as.numeric(phi[,"PC1"])))
c = as.data.frame(c)
c[order(as.numeric(c[,2]),decreasing= T),]
```
Create the new feature from the PC1 to replace 3 predictors
```{r}
datPCAbin.std <- as.data.frame(scale(datPCAbin))
#Create a new feature
dat2 <- dat #Preserving the original data frame []until this work is complete
dat2$ServCode_DRG_ICD <-  phi["ICD.PROCEDURE.CODE1",1]*datPCAbin.std$ICD.PROCEDURE.CODE1 +
phi["ICD9_DIAGNOST_CODE2",1]*datPCAbin.std$ICD9_DIAGNOST_CODE2 +      phi["HOSPITAL.SERVICE.CODEG",1]*datPCAbin.std$HOSPITAL.SERVICE.CODEG +
phi["ICD.PROCEDURE.CODE5",1]*datPCAbin.std$ICD.PROCEDURE.CODE5 +
phi["HOSPITAL.SERVICE.CODEB",1]*datPCAbin.std$HOSPITAL.SERVICE.CODEB +
phi["ICD9_DIAGNOST_CODE3",1]*datPCAbin.std$ICD9_DIAGNOST_CODE3 +
phi["PATIENT.DRG4",1]*datPCAbin.std$PATIENT.DRG4 + phi["PATIENT.DRG1",1]*datPCAbin.std$PATIENT.DRG1
summary(dat2$ServCode_DRG_ICD)

```
# Return from dat2 to dat and drop the four variables used to create the new one.

```{r}
dat <- dat2
###replace 4 PCA variable
dat$PATIENT.DRG <- NULL
dat$HOSPITAL.SERVICE.CODE <- NULL
dat$ICD.PROCEDURE.CODE <- NULL
dat$ICD9_DIAGNOST_CODE <- NULL
dat$Age<- as.factor(dat$Age)
dat <- dat[c(1:(ncol(dat)-2),ncol(dat),ncol(dat)-1)]#reorder the column, to put target in the end
summary(dat)
if(Sys.info()["user"]=="luxiong"){
  write.xlsx(dat,file="/Users/luxiong/Dropbox/Actuarial research seminar/Humana case competition/Team folder/Tingting Sun/Dr. Xiong/dat_afterPCA.xlsx")
}else{
  write.xlsx(dat,file="C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/dat_afterPCA.xlsx")
}
#If some levels are not combined correctly, can run this type of code to combine
# dat[,var] <- as.factor(dat[,var])
# var.levels <- levels(dat[,var])
# dat[,var] <- mapvalues(dat[,var],var.levels,c("F","M","F"))
```


#TASK 4 Select an interaction
```{r}
#See if there is any interation between the Patient.Sex.Code and Age
if(Sys.info()["user"]=="luxiong"){
  dat <- read.xlsx("/Users/luxiong/Dropbox/Actuarial research seminar/Humana case competition/Team folder/Tingting Sun/Dr. Xiong/dat_afterPCA.xlsx")
}else{
  dat <- read.xlsx("C:/Users/HP/Dropbox/Team folder/Tingting Sun/PA/dat_afterPCA.xlsx")
}

lapply(dat, class)#Check the class of each column
dat[,-c(ncol(dat)-1,ncol(dat))]=lapply(dat[,-c(ncol(dat)-1,ncol(dat))], factor)#Convert to factors, except last 2 columns
lapply(dat, class)

ggplot(dat,aes(x=PATIENT.SEX.CODE,y=log(gapdays+1),fill=Age))+
  geom_boxplot()+
  facet_wrap(~PATIENT.SEX.CODE,scale="free")

#See if there is any interation between the IO_Code and PATIENT.SEX.CODE
ggplot(dat,aes(x=IO_CODE,y=log(gapdays+1),fill=PATIENT.SEX.CODE))+
  geom_boxplot()+
  facet_wrap(~IO_CODE,scale="free")

#Check the count in each level to make sure there is enough observations.
nrow(subset(dat,IO_CODE=="I" & PATIENT.SEX.CODE== "F"))
nrow(subset(dat,IO_CODE=="I" & PATIENT.SEX.CODE== "M"))
nrow(subset(dat,IO_CODE=="O" & PATIENT.SEX.CODE== "F"))
nrow(subset(dat,IO_CODE=="O" & PATIENT.SEX.CODE== "M"))

#See if there is any interation between the IO_Code and AGE
ggplot(dat,aes(x=IO_CODE,y=log(gapdays+1),fill=Age))+
  geom_boxplot()+
  facet_wrap(~IO_CODE,scale="free")
```

Task 5 Select a distribution and link function
```{r}
library(caret)
set.seed(1234)
#Remove the order, because the later model not worker well with Age order
dat$Age <- factor( dat$Age , ordered = FALSE )
dat$Patient.Days <- factor( dat$Patient.Days , ordered = FALSE )
dat$income_level <- factor( dat$income_level , ordered = FALSE )

partition <- createDataPartition(dat$gapdays, list = FALSE, p = .75)
train <- dat[partition, ]
test <- dat[-partition, ]
print("TRAIN")
mean(train$gapdays)

print("TEST")
mean(test$gapdays)
GLMols <- glm(gapdays ~ ., family = gaussian(), data = train)
summary(GLMols)
print("AIC")
AIC(GLMols)
predict <- predict(GLMols,newdata=test,type="response")
print("RMSE")
sqrt(sum((test$gapdays-predict)^2)/nrow(test))
```



# This GLM uses a Gamma distribution and the log link. The interaction has been added.
```{r}
GLMgamma <- glm(gapdays+1 ~ .+ IO_CODE:PATIENT.SEX.CODE + IO_CODE:Age, family = Gamma(link = "log"), data = train)
#You can understand the above code from this Wiki:
#泊松回归假设反应变量Y是泊松分布，并假设它期望值的对数可由一组未知参数进行线性表达。
summary(GLMgamma)
print("AIC")
AIC(GLMgamma)
predict <- predict(GLMgamma,newdata=test,type="response")
print("RMSE")
sqrt(sum((test$gapdays-predict)^2)/nrow(test))
```
# GLM inverse.gaussian.
```{r}
GLMgamma <- glm(gapdays+1 ~ .+ IO_CODE:PATIENT.SEX.CODE + IO_CODE:Age, family = inverse.gaussian(link = "log"), data = train)
#You can understand the above code from this Wiki:
#泊松回归假设反应变量Y是泊松分布，并假设它期望值的对数可由一组未知参数进行线性表达。
summary(GLMgamma)
print("AIC")
AIC(GLMgamma)
predict <- predict(GLMgamma,newdata=test,type="response")
print("RMSE")
sqrt(sum((test$gapdays-predict)^2)/nrow(test))
```
# GLM gaussian
```{r}
GLMgamma <- glm(gapdays+1 ~ .+ IO_CODE:PATIENT.SEX.CODE + IO_CODE:Age, family = gaussian(link = "log"), data = train)
summary(GLMgamma)
print("AIC")
AIC(GLMgamma)
predict <- predict(GLMgamma,newdata=test,type="response")
print("RMSE")
sqrt(sum((test$gapdays-predict)^2)/nrow(test))
```
TASK 6 Select features using AIC or BIC
```{r}
library(MASS)
GLMgamma1 <- glm(gapdays+1 ~ 1, family = Gamma(link = "log"), data = train) 
#Sets up the null model for use with forward selection. The full model was obtained in Task 5.
stepAIC(GLMgamma1, direction = "forward", k = log(nrow(train)), scope = list(upper = GLMgamma, lower = GLMgamma1),print =FALSE)
```

# I next run this model to ensure everything looks right and draw any final conclusions.
```{r}
GLMgammaR <- glm(gapdays+1 ~ DOCTOR.NUMBER + ServCode_DRG_ICD + 
    Age + Surgeon + PATIENT.MARITAL.STATUS + Nur.Stat + DISCHARGE.STATUS + 
    income_level + Patient.Days + PATIENT.RACE.CODE + IO_CODE, family = Gamma(link = "log"), data = train)
summary(GLMgammaR)
print("AIC")
AIC(GLMgammaR)
predict <- predict(GLMgammaR,newdata=test,type="response")
print("RMSE")
sqrt(sum((test$Crash_Score-predict)^2)/nrow(test))
```
# Combine 3 levels to base level: DOCTOR.NUMBER=5, Surgeon=5, Patient.Days=30+
```{r}
var="DOCTOR.NUMBER"
var.levels <- levels(train[,var])
train[,var] <- mapvalues(train[,var],var.levels,c("1","2","3","4","1"))
var.levels <- levels(test[,var])
test[,var] <- mapvalues(test[,var],var.levels,c("1","2","3","4","1"))

var="Surgeon"
var.levels <- levels(train[,var])
train[,var] <- mapvalues(train[,var],var.levels,c("1","2","3","4","1"))
var.levels <- levels(test[,var])
test[,var] <- mapvalues(test[,var],var.levels,c("1","2","3","4","1"))

var="Nur.Stat"
var.levels <- levels(train[,var])
train[,var] <- mapvalues(train[,var],var.levels,c("isNA","isNA","WS"))
var.levels <- levels(test[,var])
test[,var] <- mapvalues(test[,var],var.levels,c("isNA","isNA","WS"))

var="Patient.Days"
var.levels <- levels(train[,var])
train[,var] <- mapvalues(train[,var],var.levels,c("0-19","20-30","0-19"))
var.levels <- levels(test[,var])
test[,var] <- mapvalues(test[,var],var.levels,c("0-19","20-30","0-19"))

var="PATIENT.RACE.CODE"
var.levels <- levels(train[,var])
train[,var] <- mapvalues(train[,var],var.levels,c("AIMNT","AIMNT","DH","WX"))
var.levels <- levels(test[,var])
test[,var] <- mapvalues(test[,var],var.levels,c("AIMNT","AIMNT","DH","WX"))
```

Task 7
```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

